###R markdown code##

---
title: "EE 283 - Lab 1"
author: "Kevin Cabreraaaaaaaaaaa"
date: "1/10/2019"
output: html_document
---

# Chapter 7: Tibbles
### MEMES
1). Create a small tibble manually
```{r}
tibble_example <- tibble(ranks = 1:4, countries = c('Spain', 'France', 'Ukraine', 'Sweden'),
                         age = c(23,27,28,25))
tibble_example
```
2). How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).
```{r}
head(mtcars)
tibble_example
```
*Tibbles will explicitly say that they are tibbles whereas Data Frames will just display the information as a table*

3). Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?
```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

df_tibble <- tibble(abc=1, xyz="a")
df_tibble$x
df_tibble[, "xyz"]
df_tibble[, c("abc", "xyz")]
```
*df_tribble$x returns an error rather than returning the data for column 'xyz' (there is no partial matching), df_tibble[, "xyz"] also returns the data for column xyz but returns it as another tibble, df_tibble[, c("abc", "xyz")] also returns the row data for both columns but just as a tibble again*

4). Practice referring to non-syntactic names in the following data frame by:
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```
* Extracting the variable called 1.
```{r}
Column_One <- annoying$`1`
```
* Plotting a scatterplot of 1 vs 2.
```{r}
plot(x = annoying$`1`, y = annoying$`2`)
```
* Creating a new column called 3 which is 2 divided by 1
```{r}
annoying %>% mutate(Col3 = annoying$`2`/annoying$`1`)
```
* Renaming the columns to one, two and three
```{r}
annoying %>% mutate(Col3 = annoying$`2`/annoying$`1`) %>% rename("one"=`1`, "two" = `2`, "three" = Col3)
```

5). What does tibble::enframe() do? When might you use it?
```{r}
tibble::enframe(c(a = 5, b = 7))
```
*It creates a two-column data frame from vectors or lists. It might be useful when you need to convert raw data into a useable table format*

6). What option controls how many additional column names are printed at the footer of a tibble?
```{r}
annoying %>%
  print(n = 5)
```

# Chapter 8: Data Import
1). What function would you use to read a file where fields were separated with “|”?
```{r}
read_delim(file, delim = "|")
```

2). Apart from file, skip, and comment, what other arguments do read_csv() and read_tsv() have in common?
*n_max, guess_max, progress, col_names, col_types, locale, na, quoted_na, quote, tim_ws*

4). Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By convention, read_csv() assumes that the quoting character will be ", and if you want to change it you’ll need to use read_delim() instead. What arguments do you need to specify to read the following text into a data frame?

```{r}
read_delim("x,y\n1,'a,b'", delim = ",", quote = "'")
```

5). Identify what is wrong with each of the following inline CSV files. What happens when you run the code?
```{r}
read_csv("a,b\n1,2,3\n4,5,6")
```
*The number of columns doesn't match the number of variables per row*
```{r}
read_csv("a,b,c\n1,2\n1,2,3,4")
```
*There is a missing value in the second row (replaced with NA) and there is an extra value in the second row*
```{r}
read_csv("a,b\n\"1")
```
*There is an extra backslash after the line break and there is a missing value in the row replaced by NA*
```{r}
read_csv("a,b\n1,2\na,b")
```
*Despite there being integer values within the rows, they're identified as characters*
```{r}
read_csv("a;b\n1;3")
```
*The code is recognizing the semicolon as part of the value instead of acting as a comma to show individual values*

# Chapter 9: Tidy Data
1). Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?
```{r}
table2_mod <-  spread(table2, key = type, value = count)
ggplot(table2_mod, aes(year, cases)) +
  geom_line(aes(group = country), color = "grey50") + 
  geom_point(aes(color=country))
```

2). Why does this code fail?
```{r}
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
```
*The code fails because the column names are not put in single quotes, therefore it's not being read as a name but a value*

# Chapter 10: Relational Data with dplyr
1). Add a surrogate key to flights
```{r}
flights %>% mutate(row_number()) %>% print(n=5, width=Inf)
```

2). Add the location of the origin and destination (i.e. the lat and lon) to flights.
```{r}
airports_select <- airports %>%
  select(faa, lat, lon)
airports_select %>%
  right_join(flights, by = c("faa"="origin")) %>%
  rename("origin"=faa) %>%
  rename("origin_lat"=lat) %>%
  rename("origin_lon"=lon) %>%
  left_join(airports_select, by = c("dest"="faa")) %>%
  rename("dest_lat"=lat) %>%
  rename("dest_lon"=lon)
```

